---
title: "Reproductive101"
author: "ND"
date: "2024-08-10"
output: html_document
---
# Research question:Are Sockeye salmon escapement goals being met in recent years in Bristol Bay?

## Data preperation
### Tools Loading
```{r}
library(dplyr)
library(tidyr)
library(DT)
```
### Import data
```{r}
pacific_salmon<-read.csv("/Users/mac/Library/CloudStorage/OneDrive-JohnsHopkins/R/sasap-test/MandV2016.csv")

salmon_oceanak<-read.csv("/Users/mac/Library/CloudStorage/OneDrive-JohnsHopkins/R/sasap-test/ADFG_firstAttempt_reformatted.csv")
```

## Data Cleaning and Manipulation

### Clean

1. Clean the escapement goals dataset
```{r}
#Quick Look of the data
head(pacific_salmon)
head(salmon_oceanak)
  
## Check column type and Filter to just the Bristol Bay region and the Sockeye salmon species for Pacific salmon
sockeye_goals<-pacific_salmon %>% 
  filter(Region=="Bristol Bay",Species=="Sockeye") %>% 
  mutate(Lower=as.integer(Lower),Initial.Year=as.integer(Initial.Year)) %>% 
  select(System,Lower,Upper,Initial.Year) %>% 
  drop_na()
datatable(sockeye_goals)
```



2. Clean the escapement dataset
```{r}
## Check column type and Filter to just the Bristol Bay region and the Sockeye salmon species for oceanak

sockeye_escapement<-salmon_oceanak %>% 
  filter(SASAP.Region=="Bristol Bay",Species=="Sockeye",Location %in% sockeye_goals$System) %>% 
  separate(sampleDate,c("Year","Month","Day"),sep = "-") %>% 
  group_by(Location,Year) %>% 
  summarise(Escapement=sum(DailyCount))

datatable(sockeye_escapement)
  
```
## Intergrate
Join the escapement goal lower and upper bounds onto the annual total escapement counts 
```{r}
escapement_with_goals <- 
  left_join(sockeye_escapement, sockeye_goals, by=c("Location"="System")) %>% 
  mutate(is_goal_within = ifelse(Escapement < Upper & Escapement > Lower, TRUE, FALSE),
         drop_by_year = ifelse(Year >= Initial.Year, FALSE, TRUE))
datatable(escapement_with_goals)

```
left_join(sockeye_escapement, sockeye_goals, by=c("Location"="System"))
## Graph Analyze
```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
```

### New data install
```{r}
permits <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Aa3c58bd6-481e-4c64-aa93-795df10a4664", method = "libcurl"),
                    stringsAsFactors = FALSE)
```

### Geoms/Aesthetics
```{r}
ggplot(permits, aes(Value, StdDev))+
  geom_point()
```


```{r}
ggplot(permits, aes(Value, StdDev, color = Region))+
  geom_point()
```


```{r}
# sizing
ggplot(permits, aes(Value, StdDev, color = Gear, size = Range))+
  geom_point()
```


```{r}
# histogram
ggplot(permits, aes(Value))+
  geom_histogram(binwidth = 1e4)
```


```{r}
# box plot
ggplot(permits, aes(Gear, Value))+
  geom_boxplot()
```


```{r}
# violin plot
permits_ci<- permits %>% 
  filter(Region=="Cook Inlet")

ggplot(permits_ci, aes(Gear, Value))+
  geom_violin()
```

```{r}
# multi plot
ggplot(permits, aes(Value, StdDev))+
  geom_point()+
  geom_smooth()
```


```{r}
# bar plot

permits_sum <- permits %>% 
  group_by(Gear, Region) %>% 
  summarize(mean_value = mean(Value, na.rm = T))

ggplot(permits_sum, aes(x=Gear,y=mean_value, fill=Region))+
  geom_bar(position = "stack", stat="identity")
```

```{r}
ggplot(permits_sum, aes(x = Gear, y = mean_value, fill = Region)) +
  geom_bar(position = "dodge", stat = "identity")
```
### Setting plot limits

Plot limits can be controlled one of the three ways:
- Filter the data (because limits are auto calculated from the data ranges)
- Set the limits argument on one or both scales
- Set the xlim and ylim argument in coord_cartesian()

```{r}
permits_se_seine <- permits %>% 
  filter(Gear == "Purse Seine", 
         Region == "Southeast")

ggplot(permits_se_seine, aes(Year, Value)) + 
  geom_point() +
  geom_line()+
  coord_cartesian(xlim = c(2000, max(permits_se_seine$Year)),
                  ylim = c(0, max(permits_se_seine$Value)))
  
```

### Scales

```{r}
ggplot(permits, aes(Value, StdDev, color = Gear)) + 
  geom_point()+
   scale_color_manual(values = c("red", "orange", "yellow", "green", "blue", "purple", "violet")) # ROYGBIV
```
```{r}
ggplot(permits, aes(Value, StdDev, color = Gear)) + 
  geom_point() + 
  scale_color_hue(h = c(270, 360)) # blue to red
```


```{r}
# scales
ggplot(permits, aes(Value, StdDev, color = Gear)) + 
  geom_point() + 
  scale_x_log10()
```

```{r}
permits %>%
  group_by(Gear) %>% 
  summarize(meanval = mean(Value, na.rm = TRUE)) %>% 
  ggplot(aes(Gear, meanval)) +
  geom_col() +
  scale_x_discrete(labels = sort(unique(permits$Gear)))
```

```{r}
ggplot(permits_se_seine, aes(Year, Value)) + 
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = c(1990, 2010))
```

### Facets
Facets allow us to create a powerful visualization called a small multiple:

```{r}
ggplot(permits_ci, aes(Year, Value)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ Gear)
```

```{r}
ggplot(permits_ci, aes(Year, Value)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ Gear, ncol = 1)
```


### Plot customization w/ themes

```{r}
ggplot(permits, aes(Value, StdDev, color = Gear)) + 
  geom_point() +
  theme_classic()
```



```{r}
ggplot(permits, aes(Value, StdDev, color = Gear)) + 
  geom_point() +
  theme_classic() +
  theme(legend.position = "bottom",
        legend.background = element_rect(fill = "#EEEEEE", color = "black"),
        legend.title = element_blank(),
        axis.title = element_text(size = 16))
```

```{r}
ggplot(permits, aes(Value, StdDev, color = Gear)) +
  geom_point() +
  theme_classic() +
  theme(legend.position = c(1, 1),
        legend.justification = c(1,1),
        legend.direction = "horizontal",
        legend.title = element_blank()) +
  xlab("Permit Value (unadjusted USD)") +
  ylab("Permit Std. Dev. (unadjusted USD)") +
  ggtitle("Permit Standard Deviation against Permit Value by Gear Type",
          "or: I wish I owned a seine permit")
```

### Saving plot

```{r}
ggsave("permit_stddev_vs_value.png", width = 6, height = 6)
```

